SPEI by stations

### 2017-11-08 analysis/drought/explore_drought_station.Rmd
add drought station

---
title: "SPEI by stations"
author: "AJ Perez-Luque (@ajpelu)"
date: "2016 Sep"
output:  
  md_document:
    toc: true
    variant: markdown_github
---

@@

```{r, eval=FALSE}
# Map of station

@@

```{r}
# SPEI Cadiar
spei_ca <- compute_spei(ca, rango_mensual = c(1,3,6,12,24), inicio = c(2000,9))

@@

```{r}
# SPEI Padul
spei_pa <- compute_spei(pa, rango_mensual = c(1,3,6,12,24), inicio = c(2000,9))

@@

```{r}
# SPEI Jerez 
spei_je <- compute_spei(je, rango_mensual = c(1,3,6,12,24), inicio = c(2000,9))

@@

```{r}

@@

```{r}
# We used only the 5514 station (the most long term)
# Filter by station and prec
# Filter from 1950 
st5514 <- ltdf %>% 
  filter(VARIABLE == 'PM1') %>% 
  filter(cod == '5514') %>% 
  filter(hyear >= 1950) %>% 
  dplyr::select(fecha, month, hmonth, hyear, hyear_f, pre=VALOR)

@@

st5514_pro  %>% 
  filter(!(hyear_f %in% c('2004-2005', '2011-2012'))) %>% 
  ggplot(aes(x=as.factor(hmonth), y=csum)) + 
  geom_boxplot() +  
  geom_smooth(aes(group=1), se=TRUE, level=0.95, colour='black') + 
  geom_line(data=subset(st5514_pro, hyear_f == '2004-2005'),
            aes(x=as.factor(hmonth), y=csum, group=1), colour='blue') +
  geom_line(data=subset(st5514_pro, hyear_f == '2011-2012'),
            aes(x=as.factor(hmonth), y=csum, group=1),
            colour='red') +
  theme_bw() + 
  theme(strip.background = element_rect(fill = NA, color = "black")) + 
  ylab('Precipitation (mm) acumulated') + xlab('') + 
  ggtitle('Granada Base Aerea, 1950-2015') + 
  scale_x_discrete(labels=c("1"="Sep", "2"="Oct", "3"="Nov", "4"="Dec", "5"="Jan", "6"="Feb",
                            "7"="Mar", "8"="Apr", "9"="May", "10"="Jun", "11"="Jul", "12"="Aug"))
```

@@

```{r}
# Cumulated precipitation from 2000 to 2015 for 5514 station 
st5514_2000  <- ltdf %>% 
  filter(VARIABLE == 'PM1') %>% 
  filter(cod == '5514') %>% 
  filter(hyear >= 1999) %>% 
  dplyr::select(fecha, month, hmonth, hyear, hyear_f, pre=VALOR) %>% 
  group_by(hyear_f, hmonth) %>% 
  dplyr::summarise(pre = sum(pre)) %>%
  dplyr::mutate(csum = cumsum(pre))

st5514_2000 %>% 
  filter(!(hyear_f %in% c('2004-2005', '2011-2012'))) %>% 
  ggplot(aes(x=as.factor(hmonth), y=csum)) + 
  geom_boxplot() +  
  geom_smooth(aes(group=1), se=TRUE, level=0.95, colour='black') + 
  geom_line(data=subset(st5514_2000, hyear_f == '2004-2005'),
            aes(x=as.factor(hmonth), y=csum, group=1), colour='blue') +
  geom_line(data=subset(st5514_2000, hyear_f == '2011-2012'),
            aes(x=as.factor(hmonth), y=csum, group=1),
            colour='red') +
  theme_bw() + 
  theme(strip.background = element_rect(fill = NA, color = "black")) + 
  ylab('Precipitation (mm) acumulated') +  xlab('') + 
  ggtitle('Granada Base Aerea, 2000-2015') + 
  scale_x_discrete(labels=c("1"="Sep", "2"="Oct", "3"="Nov", "4"="Dec", "5"="Jan", "6"="Feb",
                            "7"="Mar", "8"="Apr", "9"="May", "10"="Jun", "11"="Jul", "12"="Aug"))

@@

rba %>% 
  filter(!(hyear_f %in% c('2004-2005', '2011-2012'))) %>% 
  ggplot(aes(x=as.factor(hmonth), y=csum)) + 
  geom_boxplot() +  
  geom_smooth(aes(group=1), se=TRUE, level=0.95, colour='black') + 
  geom_line(data=subset(rba, hyear_f == '2004-2005'),
            aes(x=as.factor(hmonth), y=csum, group=1), colour='blue') +
  geom_line(data=subset(rba, hyear_f == '2011-2012'),
            aes(x=as.factor(hmonth), y=csum, group=1),
            colour='red') +
  theme_bw() + 
  theme(strip.background = element_rect(fill = NA, color = "black")) + 
  ylab('Precipitation (mm) acumulated') + 
  xlab('months (hydrological year; 1= sep)') 
```

@@

# PLOT 
ggplot(df, aes(x=fecha, y=value, fill=signo)) +  
  geom_bar(stat = "identity") + 
  scale_fill_manual(values = c("pos" = "darkblue", "neg" = "red")) + 
  facet_wrap(~spei_ord, ncol=1) +
  scale_x_date(date_breaks = "10 year", date_labels = "%Y") + 
  theme_bw() + 
  ylab('SPEI') + 
  xlab('year') + 
  theme(legend.position = "none", 
        strip.background = element_blank()) +
  ggtitle(site) + 
  geom_vline(xintercept=as.numeric(as.Date("2005-01-01"))) +
  geom_vline(xintercept=as.numeric(as.Date("2005-12-31")))

>>>>

---
title: "SPEI by stations"
author: "AJ Perez-Luque (@ajpelu)"
date: "2016 Sep"
output:  
  md_document:
    toc: true
    variant: markdown_github

---

@@

```{r eval=FALSE}
# Map of station

@@

```{r speiCadiar}
# SPEI Cadiar
spei_ca <- compute_spei(ca, rango_mensual = c(1,3,6,12,24), inicio = c(2000,9))

@@

```{r speiPadul}
# SPEI Padul
spei_pa <- compute_spei(pa, rango_mensual = c(1,3,6,12,24), inicio = c(2000,9))

@@

```{r speiJerez}
# SPEI Jerez 
spei_je <- compute_spei(je, rango_mensual = c(1,3,6,12,24), inicio = c(2000,9))

@@

```{r prec_curvesRIA}

@@

```{r prec_curvesBaseAerea}
# We used only the 5514 station (the most long term)
# Filter by station and prec
# Filter from 1950 
st5514 <- ltdf %>% 
  filter(VARIABLE == 'PM1') %>% 
  filter(cod == '5514') %>% 
  filter(hyear >= 1950) %>% 
  dplyr::select(fecha, month, hmonth, hyear, hyear_f, pre=VALOR)

@@

st5514_pro  %>% 
  filter(!(hyear_f %in% c('2004-2005', '2011-2012'))) %>% 
  ggplot(aes(x=as.factor(hmonth), y=csum)) + 
  geom_boxplot() +  
  geom_smooth(aes(group=1), se=TRUE, level=0.95, colour='black') + 
  geom_line(data=subset(st5514_pro, hyear_f == '2004-2005'),
            aes(x=as.factor(hmonth), y=csum, group=1), colour='blue') +
  geom_line(data=subset(st5514_pro, hyear_f == '2011-2012'),
            aes(x=as.factor(hmonth), y=csum, group=1),
            colour='red') +
  theme_bw() + 
  theme(strip.background = element_rect(fill = NA, color = "black")) + 
  ylab('Precipitation (mm) acumulated') + xlab('') + 
  ggtitle('Granada Base Aerea, 1950-2015') + 
  scale_x_discrete(labels=c("1"="Sep", "2"="Oct", "3"="Nov", "4"="Dec", "5"="Jan", "6"="Feb",
                            "7"="Mar", "8"="Apr", "9"="May", "10"="Jun", "11"="Jul", "12"="Aug")) +
  annotate("text", label = "2004-2005", x = 2, y = 500, size = 6, colour = "blue") + 
  annotate("text", label = "2011-2012", x = 2, y = 550, size = 6, colour = "red") + 
  annotate("text", label = "mean 1950-2015", x = 3, y = 600, size = 6, colour = "black")
```

@@

```{r prec_curvesBaseAerea2000}
# Cumulated precipitation from 2000 to 2015 for 5514 station 
st5514_2000  <- ltdf %>% 
  filter(VARIABLE == 'PM1') %>% 
  filter(cod == '5514') %>% 
  filter(hyear >= 1999) %>% 
  dplyr::select(fecha, month, hmonth, hyear, hyear_f, pre=VALOR) %>% 
  group_by(hyear_f, hmonth) %>% 
  dplyr::summarise(pre = sum(pre)) %>%
  dplyr::mutate(csum = cumsum(pre))

st5514_2000 %>% 
  filter(!(hyear_f %in% c('2004-2005', '2011-2012'))) %>% 
  ggplot(aes(x=as.factor(hmonth), y=csum)) + 
  geom_boxplot() +  
  geom_smooth(aes(group=1), se=TRUE, level=0.95, colour='black') + 
  geom_line(data=subset(st5514_2000, hyear_f == '2004-2005'),
            aes(x=as.factor(hmonth), y=csum, group=1), colour='blue') +
  geom_line(data=subset(st5514_2000, hyear_f == '2011-2012'),
            aes(x=as.factor(hmonth), y=csum, group=1),
            colour='red') +
  theme_bw() + 
  theme(strip.background = element_rect(fill = NA, color = "black")) + 
  ylab('Precipitation (mm) acumulated') +  xlab('') + 
  ggtitle('Granada Base Aerea, 2000-2015') + 
  scale_x_discrete(labels=c("1"="Sep", "2"="Oct", "3"="Nov", "4"="Dec", "5"="Jan", "6"="Feb",
                            "7"="Mar", "8"="Apr", "9"="May", "10"="Jun", "11"="Jul", "12"="Aug")) +
  annotate("text", label = "2004-2005", x = 2, y = 500, size = 6, colour = "blue") + 
  annotate("text", label = "2011-2012", x = 2, y = 550, size = 6, colour = "red") + 
  annotate("text", label = "mean 2000-2015", x = 3, y = 600, size = 6, colour = "black")

@@

rba %>% 
  filter(!(hyear_f %in% c('2004-2005', '2011-2012'))) %>% 
  ggplot(aes(x=as.factor(hmonth), y=csum)) + 
  geom_boxplot() +  
  geom_smooth(aes(group=1), se=TRUE, level=0.95, colour='black') + 
  geom_line(data=subset(rba, hyear_f == '2004-2005'),
            aes(x=as.factor(hmonth), y=csum, group=1), colour='blue') +
  geom_line(data=subset(rba, hyear_f == '2011-2012'),
            aes(x=as.factor(hmonth), y=csum, group=1),
            colour='red') +
  theme_bw() + 
  theme(strip.background = element_rect(fill = NA, color = "black")) + 
  ylab('Precipitation (mm) acumulated') + 
  xlab('months (hydrological year; 1= sep)') +
  annotate("text", label = "2004-2005", x = 2, y = 500, size = 6, colour = "blue") + 
  annotate("text", label = "2011-2012", x = 2, y = 550, size = 6, colour = "red") + 
  annotate("text", label = "mean 1950-2009", x = 3, y = 600, size = 6, colour = "black")
```

@@

# PLOT 
ggplot(df, aes(x=fecha, y=value, fill=signo)) +  
  geom_bar(stat = "identity") + 
  scale_fill_manual(values = c("pos" = "darkblue", "neg" = "red")) + 
  facet_wrap(~spei_ord, ncol=1) +
  scale_x_date(date_breaks = "10 year", date_labels = "%Y") + 
  theme_bw() + 
  ylab('SPEI') + 
  xlab('year') + 
  theme(legend.position = "none", 
        strip.background = element_blank()) +
  ggtitle(site) + 
  geom_vline(xintercept=as.numeric(as.Date("2005-01-01"))) +
  geom_vline(xintercept=as.numeric(as.Date("2005-12-31"))) +
  scale_x_date(date_breaks = "2 year", date_labels = "%Y")
